---
title: "Random sample"
author: "Sara Colom"
date: "`r Sys.Date()`"
output: 
  github_document:
    df_print: kable
---

# Read libraries and data

```{r, message = F}
library(dplyr)
library(tidyr)
library(stringr)

path_to_proj <- Sys.getenv("HOME_SAFE")

dat <- readxl::read_excel(paste(path_to_proj, "/Quantitative/data/report_1_data/home_safe_data_upload_090623.xlsx", sep = "")) %>% 
  janitor::clean_names()
```


# Random sample

Group by agency and pull 15% of case numbers w/o replacement.

```{r}
dat %>% 
  count(reporting_agency) %>% 
  arrange(n) %>% 
  head()
```

Some reporting agencies have less than 10 cases.

Assign agency as a `small_agency` (10 cases or less) or `not_small` (more than 10 cases)

```{r, message = FALSE}
dat <- dat %>% 
  add_count(reporting_agency) %>% 
  mutate(`agency_size` = if_else(n <= 10, "small_agency", "not_small"))

small_agencies <- dat %>% 
  filter(agency_size == "small_agency") %>% 
  droplevels()

all_other_agencies <- dat %>% 
  anti_join(small_agencies) %>% 
  droplevels()
```

# Alternative I.

If N < 10 capture `10%  across all SMALL agencies or just capture 1 per agency 
Anything more we capture `10%`

```{r}
set.seed(123)

random_sample_small_agencies_i <- small_agencies %>% 
  group_by(reporting_agency) %>% 
  sample_n(1)
```


# Alternative II

We pool all those with 10 or less cases as one giant 'small agency' pool, and select 10%, then for those > 10 cases randomly pool 10%.

```{r}
sample_n <- nrow(small_agencies)*.1 %>% 
  round(1)

set.seed(123)

random_sample_small_agencies_ii <- small_agencies %>% 
  sample_n(sample_n)
```


# Sample all other agencies

```{r}
all_other_agencies <- all_other_agencies %>% 
  mutate(sample_n = ceiling(n*0.1))

set.seed(123)

random_sample_other_agencies <- all_other_agencies %>% 
  group_by(reporting_agency) %>% 
  group_modify(~sample_n(.x, sample_n[1])) %>% 
  ungroup()
```


# Distribution of samples

The smaller agencies

```{r}
dist_small <- small_agencies %>% 
  ungroup() %>% 
  select(race_1, ethnicity, reporting_agency) %>% 
  pivot_longer(everything(), names_to = "variable") %>% 
  group_by(variable, value) %>% 
  summarise(n = n()) %>% 
  ungroup(value) %>% 
  mutate(statistic = n/sum(n)) %>% 
  bind_rows(
    small_agencies %>% 
    summarise(variable = "age",
              sd = sd(age, na.rm = T),
              statistic = mean(age, na.rm = T))
  ) %>% 
  mutate(across(where(is.numeric), ~round(.x, 2))) %>% 
  mutate(sample = "Full Sample") %>% 
  mutate(statistic = case_when(variable == "age" ~ paste(statistic, " (", sd, ")", sep = ""),
                               is.na(statistic) ~ NA,
                               TRUE ~ paste(n, " ", "(", statistic*100, "%",")", sep = ""))) %>% 
  select(variable, value, statistic, sample)
  

dist_small_i <- random_sample_small_agencies_i %>% 
  select(race_1, ethnicity, reporting_agency) %>% 
  pivot_longer(everything(), names_to = "variable") %>% 
  group_by(variable, value) %>% 
  summarise(n = n()) %>% 
  ungroup(value) %>% 
  mutate(statistic = n/sum(n)) %>% 
  bind_rows(
    small_agencies %>% 
    summarise(variable = "age",
              sd = sd(age, na.rm = T),
              statistic = mean(age, na.rm = T))
  ) %>% 
  mutate(across(where(is.numeric), ~round(.x, 2))) %>% 
  mutate(sample = "Random Sample I") %>% 
  mutate(statistic = case_when(variable == "age" ~ paste(statistic, " (", sd, ")", sep = ""),
                               is.na(statistic) ~ NA,
                               TRUE ~ paste(n, " ", "(", statistic*100, "%",")", sep = ""))) %>% 
  select(variable, value, statistic, sample)


dist_small_ii <- random_sample_small_agencies_ii %>% 
  select(race_1, ethnicity, reporting_agency) %>% 
  pivot_longer(everything(), names_to = "variable") %>% 
  group_by(variable, value) %>% 
  summarise(n = n()) %>% 
  ungroup(value) %>% 
  mutate(statistic = n/sum(n)) %>% 
  bind_rows(
    small_agencies %>% 
    summarise(variable = "age",
              sd = sd(age, na.rm = T),
              statistic = mean(age, na.rm = T))
  ) %>% 
  mutate(across(where(is.numeric), ~round(.x, 2))) %>% 
  mutate(sample = "Random Sample II") %>% 
  mutate(statistic = case_when(variable == "age" ~ paste(statistic, " (", sd, ")", sep = ""),
                               is.na(statistic) ~ NA,
                               TRUE ~ paste(n, " ", "(", statistic*100, "%",")", sep = ""))) %>% 
  select(variable, value, statistic, sample)

full_stats_small_agency <- dist_small %>% 
  bind_rows(dist_small_i) %>% 
  bind_rows(dist_small_ii) %>% 
  pivot_wider(names_from = sample, values_from = statistic)

full_stats_small_agency
```

All other agencies

```{r}
dist_other <- all_other_agencies %>% 
  ungroup() %>% 
  select(race_1, ethnicity, reporting_agency) %>% 
  pivot_longer(everything(), names_to = "variable") %>% 
  group_by(variable, value) %>% 
  summarise(n = n()) %>% 
  ungroup(value) %>% 
  mutate(statistic = n/sum(n)) %>% 
  bind_rows(
    all_other_agencies %>% 
    summarise(variable = "age",
              sd = sd(age, na.rm = T),
              statistic = mean(age, na.rm = T))
  ) %>% 
  mutate(across(where(is.numeric), ~round(.x, 2))) %>% 
  mutate(sample = "Full Sample") %>% 
  mutate(statistic = case_when(variable == "age" ~ paste(statistic, " (", sd, ")", sep = ""),
                               is.na(statistic) ~ NA,
                               TRUE ~ paste(n, " ", "(", statistic*100, "%",")", sep = ""))) %>% 
  select(variable, value, statistic, sample)
  

dist_other_sample <- random_sample_other_agencies %>% 
  select(race_1, ethnicity, reporting_agency) %>% 
  pivot_longer(everything(), names_to = "variable") %>% 
  group_by(variable, value) %>% 
  summarise(n = n()) %>% 
  ungroup(value) %>% 
  mutate(statistic = n/sum(n)) %>% 
  bind_rows(
    random_sample_other_agencies %>% 
    summarise(variable = "age",
              sd = sd(age, na.rm = T),
              statistic = mean(age, na.rm = T))
  ) %>% 
  mutate(across(where(is.numeric), ~round(.x, 2))) %>% 
  mutate(sample = "Random Sample I") %>% 
  mutate(statistic = case_when(variable == "age" ~ paste(statistic, " (", sd, ")", sep = ""),
                               is.na(statistic) ~ NA,
                               TRUE ~ paste(n, " ", "(", statistic*100, "%",")", sep = ""))) %>% 
  select(variable, value, statistic, sample)


full_stats_other_agency <- dist_other %>% 
  bind_rows(dist_other_sample) %>% 
  pivot_wider(names_from = sample, values_from = statistic)

full_stats_other_agency
```

# Total N sampled

```{r}
random_sample_other_agencies %>% 
  nrow()

random_sample_small_agencies_i %>% 
  nrow()

random_sample_small_agencies_ii %>% 
  nrow()
```


# Export samples and stats

```{r}
output <- list(random_sample_small_agencies_i = random_sample_small_agencies_i, random_sample_small_agencies_ii = random_sample_small_agencies_ii,
               random_sample_other_agencies = random_sample_other_agencies, stats_small_agency = full_stats_small_agency, stats_all_other_agency = full_stats_other_agency)

writexl::write_xlsx(output, paste(path_to_proj, "/Quantitative/data/random_sample_iteration_1/randomized_sample.xlsx", sep = ""))

write.csv(dat, paste(path_to_proj, "/Quantitative/data/random_sample_iteration_1/og_file_used.csv", sep = ""))
```

