---
title: "Cleaning Demographic Data"
author: "Sara Colom"
date: "`r Sys.Date()`"
output: 
  github_document:
    df_print: kable
---


# Read in data & libraries

```{r, Read in Data & Libraries, warning = F, message = F}
source("R/helpers.R")
```

```{r}
demo_dat %>% 
  glimpse()
```


# Clean up demographic data

Make all variables of `character` lower case, trim trailing and starting white spaces and 'extra' inner spaces.

```{r}
demo_dat <- demo_dat %>% 
  mutate(across(where(is.character), clean_characters))

```


## Location of Participation:

remove " ca" ", ca" and "california" from the end of responses. remove special characters like ":" before laguna beach. Combine and correct misspellings. Leaving combined cities or ones that seem similar (i.e. big bear and big bear valley) alone for now, also not verifying that all of these are in fact cities (I only verified for ones I suspected having misspellings). Remove Zip codes from the end of cities listed. Combining all unknown's into one (data not collected, refused, doesn't know etc). 

### Questions:
  ca - perhaps best to change this to unknown?
  "cal city", "california city" - do we think these are the same? (Made the change in the recode for now)
  responses i.e. "coulterville/la grange" (so combo of two cities) - leave as is?
  "north side of parking lot" - ?
  "sin city" - might be sun city? (there is no sin city in california it seems)
  "ylp" - ?
  what to do with zipcodes? (shall I go look up the cities for those codes?)
  home, homeless, female, male - are in the data so double check after new data comes in
```{r}
demo_dat <- demo_dat %>% 
  mutate(location_of_participation_recode = case_when(str_ends(location_of_participation, " ca") ~ str_replace_all(location_of_participation, "ca|, ca", ""),
                                                      str_ends(location_of_participation, "california") ~ str_replace_all(location_of_participation, " california", ""),
                                                      str_detect(location_of_participation, ":") ~ "laguna beach",
                                                      str_detect(location_of_participation, "bakersf|bakerf") ~ "bakersfield",
                                                      str_detect(location_of_participation, "bartso") ~ "barstow",
                                                      str_detect(location_of_participation, "beam|beum") ~ "beaumont",
                                                      str_detect(location_of_participation, "cal city") ~ "california city",
                                                      str_detect(location_of_participation, "cailm") ~ "calimesa",
                                                      str_detect(location_of_participation, "catherd") ~ "cathedral city",
                                                      str_detect(location_of_participation, "cathey") ~ "catheys valley",
                                                      str_detect(location_of_participation, "clairem") ~ "claremont",
                                                      str_detect(location_of_participation, "cochell") ~ "coachella", 
                                                      str_detect(location_of_participation, "desesrt|desrt") ~ "desert hot springs",
                                                      str_detect(location_of_participation, "frasier") ~ "frazier park",
                                                      str_detect(location_of_participation, "idlewild") ~ "idyllwild",
                                                      str_detect(location_of_participation, "jurupa") ~"jurupa valley",
                                                      str_detect(location_of_participation, "mills hospital in b") ~ "burlingame",
                                                      str_detect(location_of_participation, "mountainview") ~"mountain view",
                                                      str_detect(location_of_participation, "murreta") ~ "murrieta",
                                                      str_detect(location_of_participation, "pachecc") ~ "pacheco",
                                                      str_detect(location_of_participation, "palms sp|palm spr|palm dpr") ~ "palm springs",
                                                      str_detect(location_of_participation, "ranch mira") ~ "rancho mirage",
                                                      str_detect(location_of_participation, "rancho cucam") ~ "rancho cucamonga",
                                                      str_detect(location_of_participation, "richomo") ~ "richmond",
                                                      str_detect(location_of_participation, "ridegc") ~ "ridgecrest",
                                                      str_detect(location_of_participation, "riverisd|riveside") ~ "riverside",
                                                      str_detect(location_of_participation, "rough &|rough and re") ~ "rough and ready",
                                                      str_detect(location_of_participation, "sacrment|sacarament|sarcamen") ~ "sacramento",
                                                      str_detect(location_of_participation, "san berna|san bera|san bernr") ~ "san bernardino",
                                                      str_detect(location_of_participation, "san dieg") ~ "san diego",
                                                      str_detect(location_of_participation, "san fra|sf") ~ "san francisco",
                                                      str_detect(location_of_participation, "sugar l") ~ "sugarloaf",
                                                      str_detect(location_of_participation, "tehac") ~ "tehachapi",
                                                      str_detect(location_of_participation, "twenty|29 pal") ~ "twentynine palms",
                                                      str_detect(location_of_participation, "vict") ~ "victorville",
                                                      str_detect(location_of_participation, "weaverv") ~ "weaverville",
                                                      str_detect(location_of_participation, "wildom") ~ "wildomar", 
                                                      str_detect(location_of_participation, "yucca") ~ "yucca valley",
                                                      # the following had zipcodes at the end of the string
                                                      str_detect(location_of_participation, "indio") ~ "indio",
                                                      str_detect(location_of_participation, "inglewood") ~ "inglewood",
                                                      str_detect(location_of_participation, "marysvil") ~ "marysville",
                                                      str_detect(location_of_participation, "turlock") ~ "turlock",
                                                      str_detect(location_of_participation, "yuba") ~ "yuba city",
                                                      str_detect(location_of_participation, "yucaipa") ~ "yucaipa",
                                                      
                                                      str_detect(location_of_participation, "data not col|doesn't know|refused|missing|unk") ~ "unknown",
                                                      is.na(location_of_participation) ~ "unknown",
                                                      TRUE ~ location_of_participation))
```


Check on location of participation
```{r}
demo_dat %>% 
  count(location_of_participation, location_of_participation_recode)
```

For just a check on cutting out ca:
```{r}
demo_dat %>% 
  filter(str_detect(location_of_participation, ", ca| ca|,|california")) %>% 
  count(location_of_participation, location_of_participation_recode)
```



Race_1 - replace american indian/alaskan native with american indian/alaskan native/indigenous. Replace asian with asian/asian american. Replace black/african american with black/african american/african. Replace pacific islander/native hawaiian with native hawaiian/pacific islander. Replace whte with white. Change all unknowns, including data not collected and n/a to "unknown".
```{r}
demo_dat <- demo_dat %>% 
  mutate(race_1_recode = case_when(str_detect(race_1,"americ") ~ "american indian/alaskan native/indigenous", 
                                      str_starts(race_1, "asian") ~ "asian/asian american",
                                      str_starts(race_1, "bla") ~ "black/african american/african",
                                      str_starts(race_1, "pacific i") ~ "native hawaiian/pacific islander",
                                      str_starts(race_1, "wh") ~ "white",
                                   str_detect(race_1, "unkn|#|client|collected") ~ "unknown",
                                   is.na(race_1) ~ "unknown",
                                      
                                         TRUE ~ race_1))
```

check
```{r}
demo_dat %>% 
  count(race_1, race_1_recode)
```

race_2 - doing the same as for race_1, and replacing no applicable with not applicable, updating "unknown". For those who put an ethnicity response in race_2, replacing with "unknown" and updating those in the ethnicity variable. 

```{r}
demo_dat <- demo_dat %>% 
  mutate(race_2_recode = case_when(str_starts(race_2,"americ") ~ "american indian/alaskan native/indigenous", 
                                      str_starts(race_2, "asian") ~ "asian/asian american",
                                      str_starts(race_2, "bla") ~ "black/african american/african",
                                      str_starts(race_2, "pacific i") ~ "native hawaiian/pacific islander",
                                      str_starts(race_2, "wh") ~ "white",
                                   str_detect(race_2, "hisp") ~ "unknown",
                                   str_detect(race_2, "unknown|unable|refused|not collected|applicable|verified|#|client|mex") ~ "unknown",
                                   is.na(race_2) ~ "unknown",
                                      
                                         TRUE ~ race_2))
```

check for race_2
```{r}
demo_dat %>% 
  count(race_2, race_2_recode)
```


Ethnicity - replace all instances of different spellings or orders of hispanic/latino/a/x to hispanic/latinx. Changing mexican/chicano, cuban, other hispanic/latino, and puerto rican to hispanic/latinx. Also, replace all instances of different spellings or orders of not hispanic/latino to non-hispanic/latinx 

```{r}
demo_dat <- demo_dat %>% 
  mutate(ethnicity_recode = case_when(str_starts(ethnicity, "no") ~ "non-hispanic/latinx",
                                      str_detect(ethnicity,"hispa") ~ "hispanic/latinx", 
                                      str_detect(race_2, "not hispanic|no hispanic") ~ "non-hispanic/latinx",
                                      str_detect(race_2, "hispanic") ~ "hispanic/latinx",
                                      str_detect(race_2, "mex") ~ "hispanic/latinx",
                                      str_detect(ethnicity, "mex|puerto|cuban") ~ "hispanic/latinx",
                                      str_detect(ethnicity, "unknown|#|applicable|data not|client|missing|rent") ~ "unknown",
                                      is.na(ethnicity) ~ "unknown",
                                      
                                      
                                         TRUE ~ ethnicity))
```

check for ethnicity 
```{r}
demo_dat %>% 
  count(ethnicity, race_2, ethnicity_recode) 
```

current_martial_status - replace unknowns and data not collected with "unknown". 

### Question: How would we like to recode domestic partnership? (Lump with married?) What to do with single? (100/103 records with single are from LA) ALSO, what to do with "no"? (both from Merced, might get fixed when column shift issue is solved)
```{r}
demo_dat <- demo_dat %>% 
  mutate(current_marital_recode = case_when(str_detect(current_marital_status, "unknown|#|applicable|data not") ~ "unknown",
                                            str_detect(current_marital_status, "domestic|single") ~ "flag",
                                      is.na(current_marital_status) ~ "unknown",
                                         TRUE ~ current_marital_status))
```

check for marital status
```{r}
demo_dat %>% 
  count(current_marital_status, current_marital_recode)
```

sexual_orientation - changed straight to "straight/heterosexual" to match CDSS variable. Changed all unknown or declined or missing responses to "unknown"

### Question: do we want to lump "another sexual orientation" or "other" into something? Perhaps an "other sexual orientation" bucket?
```{r}
demo_dat <- demo_dat %>% 
  mutate(sexual_orientation_recode = case_when(str_detect(sexual_orientation, "strai") ~ "straight/heterosexual",
                                               str_detect(sexual_orientation, "unknown|#|client|decline|data not") ~ "unknown",
                                               str_detect(sexual_orientation, "other") ~ "flag",
                                      is.na(sexual_orientation) ~ "unknown",
                                         TRUE ~ sexual_orientation))
```

check for sexual orientation
```{r}
demo_dat %>% 
  count(sexual_orientation, sexual_orientation_recode)
```

preferred_language - leaving alone the responses that don't appear to be languages for now. Combining the mandarin/cantonese responses. Changing unknowns and missings to "unknown". 
```{r}
demo_dat <- demo_dat %>% 
  mutate(preferred_language_recode = case_when(str_detect(preferred_language, "mandarin") ~ "mandarin/cantonese",
                                               str_detect(preferred_language, "unknown|#|data not") ~ "unknown",
                                               is.na(preferred_language) ~ "unknown",
                                               TRUE ~ preferred_language))
```

check for preferred language
```{r}
demo_dat %>% 
  count(preferred_language, preferred_language_recode)
```

veteran_status - combining client refused, unknown, missing, unable to verify, verified by staff, and "b" into "unknown". Leaving the response "none" for now to see if it's a column shift issue, will revisit. 

### Question: I'm guessing "none" should be lumped with "no"?
```{r}
demo_dat <- demo_dat %>% 
  mutate(veteran_status_recode = case_when(str_detect(veteran_status, "client|unknown|#|verif|b|data not|n/a") ~ "unknown",
                                           is.na(veteran_status) ~ "unknown",
                                           TRUE ~ veteran_status))
```

check for veteran status
```{r}
demo_dat %>% 
  count(veteran_status, veteran_status_recode)
```

medi_cal - combining the unknowns into one "unknown"
```{r}
demo_dat <- demo_dat %>% 
  mutate(medi_cal_recode = case_when(str_detect(medi_cal, "client|unknown|#|data not") ~ "unknown",
                                           is.na(medi_cal) ~ "unknown",
                                           TRUE ~ medi_cal))
```

check medi-cal 
```{r}
demo_dat %>% 
  count(medi_cal, medi_cal_recode)
```


doing the same for medicare
```{r}
demo_dat <- demo_dat %>% 
  mutate(medicare_recode = case_when(str_detect(medicare, "client|unknown|#|data not|n/a") ~ "unknown",
                                           is.na(medicare) ~ "unknown",
                                           TRUE ~ medicare))
```

check for medicare 
```{r}
demo_dat %>% 
  count(medicare, medicare_recode)
```

representative payee or conservator: data is real messy, seems to have many values that shouldn't be in this column. Will hold off until new dataset comes in to assess how much is a column shift issue. 
```{r}
demo_dat %>% 
  count(representative_payee_or_conservator)
```

living situation upon entry - changing the numbers to unknown for now, I'm guessing many of them are rent amount, but will wait for column shift fix to verify. Matching responses with dropdown options: combining all "homeless" responses into "homeless." Recoding "hotel" as temporary housing. Changing rent by client, rental by client, etc and "lease holder" into "rent leaseholder". Recoding descriptions of owner, whether living alone or with others, as "owner" (excluding the records where it says rental by owner).Changing "programpermanent-residential program" to "permanent-residential program" as that seems like a typo. Recoding "hospital facility" to "other". Combining unknown, n/a, data not collected as unknown. 

###Questions for the group:
    "hotel with rights" - I'm thinking this shouldn't be temporary housing, what do you all think? 
    "with others no rent"/"with others" - Is this akin to "staying with friends/family on a temp basis (temporary housing), is it "other permanent housing"?
    "board and care facility"/"residential care facility"/SNF - in the dropdown, these are in both temporary- residential program and permanent- residential program options. 
    "rent by owner" - is this rent leaseholder? (that is where I put it for now)
    "shared housing"/"living in shared house"/"living alone" - rent leaseholder? other permanent housing? Could also be owner I suppose?
    "living with family"/"living with relative" - this could be either temporary housing or other permanent housing..
    "with others rent" - my guess is this should go under "other permanent housing" which includes in () "renting a room without a lease? 
    
```{r}
demo_dat <- demo_dat %>% 
  mutate(living_sit_entry_recode = case_when(str_detect(living_situation_upon_entry, "1|2|3|4|5|6|7|8|9|0") ~ "unknown",
                                            str_detect(living_situation_upon_entry, "homeless|homless|unsheltered") ~ "homeless",
                                             str_detect(living_situation_upon_entry, "hotel") ~ "temporary housing",
                                             str_detect(living_situation_upon_entry, "rent by|rental by|lease holder") ~ "rent leaseholder",
                                             str_detect(living_situation_upon_entry, "owned|owner") ~ "owner",
                                             str_detect(living_situation_upon_entry, "programperm") ~ "permanent- residential program",
                                             str_detect(living_situation_upon_entry, "hospital") ~ "other",
                                            str_detect(living_situation_upon_entry, "hotel with rights|facility|shared hous|living with|live alone|with others") ~ "flag",
                                             str_detect(living_situation_upon_entry, "data not|unknown") ~ "unknown",
                                             is.na(living_situation_upon_entry) ~ "unknown",
                                          TRUE ~ living_situation_upon_entry))
```


check for living situation upon entry
```{r}
demo_dat %>% 
  count(living_situation_upon_entry, living_sit_entry_recode)
```

And! for a breakdown of which counties these hard to classify ones are coming up (it's many different ones it seems):
```{r}
demo_dat %>% 
  filter(str_detect(living_sit_entry_recode, "flag")) %>% 
  count(living_situation_upon_entry, reporting_agency)
```

client homeless within the last 3 years - answers should be yes or no. Leaving numbers alone for now. Changing 3 years or longer, within the last year, within the last 3 years, currently homeless to "yes". Changing client was not homeless to "no". Combining doesn't know, refused, data not collected, unknown, blank as "unknown".

Question: one response that is "n" - I'm guessing this might be no? But there is another response with n/a, which is what they might've been going for? (changed to no for now)
```{r}
demo_dat <- demo_dat %>% 
  mutate(client_homeless_last_three_years_recode = case_when(str_detect(client_homeless_within_the_last_three_years, "three years or|within the last|currently homeles") ~ "yes",
                                                  str_detect(client_homeless_within_the_last_three_years, "client was not homeles") ~ "no",
                                                  str_equal(client_homeless_within_the_last_three_years, "n") ~ "no",
                                                  str_detect(client_homeless_within_the_last_three_years, "doesn't know|refused|n/a|data not") ~ "unknown",
                                                  is.na(client_homeless_within_the_last_three_years) ~ "unknown",
                                                  TRUE ~ client_homeless_within_the_last_three_years))
```

check for homeless in last 3 years
```{r}
demo_dat %>% 
  count(client_homeless_within_the_last_three_years, client_homeless_last_three_years_recode)
```

Number of times homeless in the past 3 years - changing 0 to "client was not homeless", 1 to "one time", 2 to "two times", 3 to "three times" 4 and 4+ to "four or more times". Changing any "0-x" responses to unknown, since it seems the exact number is not known.. Changing refused, doesn't know, n/a, blank to "unknown". 

### Question (Kenny, Angelica, Sara): 
  99 is a response with 1366 answers, I'm wondering if this is the same as unknown? (like people were just using a very big number as a placeholder?) Or just a very large number of times someone (many people) were unhoused? (Putting in unknown for now)
  0-3 - where to put this one? (putting in unknown for now)
  "currently homeless" - perhaps should just change to unknown? (unsure also if this is a column shift issue, same with "yes" and "no", leaving alone for now)
  
```{r}
demo_dat <- demo_dat %>% 
  mutate(number_times_homeless_recode = case_when(str_equal(number_of_times_homelessness_occurred_in_the_last_three_years, "0") ~ "client was not homeless",
                                                  str_detect(number_of_times_homelessness_occurred_in_the_last_three_years, "0-|doesn't know|refused|data not|free form te|99") ~ "unknown",
                                                  str_equal(number_of_times_homelessness_occurred_in_the_last_three_years, "1") ~ "one time",
                                                  str_equal(number_of_times_homelessness_occurred_in_the_last_three_years, "2") ~ "two times",
                                                  str_equal(number_of_times_homelessness_occurred_in_the_last_three_years, "3") ~ "three times",
                                                  str_detect(number_of_times_homelessness_occurred_in_the_last_three_years, "0|1|2|3|4|5|6|7|8|9|four") ~ "four or more times",
                                                  is.na(number_of_times_homelessness_occurred_in_the_last_three_years) ~ "unknown",
                                                  TRUE ~ number_of_times_homelessness_occurred_in_the_last_three_years))
```

check for number of times homeless in last 3 years
```{r}
demo_dat %>% 
  count(number_of_times_homelessness_occurred_in_the_last_three_years, number_times_homeless_recode)
```

Total duration of homelessness (in last 3 years) - Changing "0" to "client was not homeless". Leaving "yes" and "no" responses for now, as I think they're a column shift issue. Combining unknown responses.Changing responses that are more than a year to all be "more than a year". changing "n" to "unknown".


### Questions:
  99 - same question as number of times homeless
  responses like "1", "2" "24" - no units.. so not sure. My guess would be those are the number of months? (left as is for now)
  "within the last 3 years" - can't really take a guess at duration, besides saying it's not more than 3 years?
  "six months to one year" "7 months to one year" "less than 12 months" "two to six months" - what to do?
  "within the last year" "within the last 3 years" "currently homeless" - seems a response to the following question (last period of homelessness). Looking at what those records have for last period, both don't really make sense, so double check after column issue fix?
  (Sara) - there must be a better way to code replacing numbers with the number written out? (i.e. 5 months --> five months)
  
```{r}
demo_dat <- demo_dat %>% 
  mutate(total_duration_homeless_recode = case_when(str_equal(total_duration_of_homelessness, "0") ~ "client was not homeless",
                                                    str_detect(total_duration_of_homelessness, "not homeless") ~ "client was not homeless",
                                                    str_detect(total_duration_of_homelessness, "blank|doesn't know|data no|refused|none|n/a|99") ~ "unknown",
                                                    str_detect(total_duration_of_homelessness, "2 year|3 year|6 year|few year|more than a year|years or longer") ~ "more than a year",
                                                    str_detect(total_duration_of_homelessness, "1 day to one month") ~"one day to one month",
                                                    str_detect(total_duration_of_homelessness, "5 mon") ~ "five months",
                                                    str_detect(total_duration_of_homelessness, "6 mon") ~ "six months",
                                                    str_detect(total_duration_of_homelessness, "to one year|less than 12 months|within the last|currently homeless") ~ "flag",
                                                    str_equal(total_duration_of_homelessness, "n") ~ "unknown",
                                                    is.na(total_duration_of_homelessness) ~ "unknown",
                                                    
                                                    TRUE ~ total_duration_of_homelessness))
```

Check for total duration of homelessness in last 3 years
```{r}
demo_dat %>% 
  count(total_duration_of_homelessness, total_duration_homeless_recode)
```

Current eviction or foreclosure - combining unknown responses, and making "none" "no". Leaving the rest for now. 

### Question:
  is 99 unknown here as well? (this one isn't a drop down menu, so wondering if I should treat the same)
```{r}
demo_dat <- demo_dat %>% 
  mutate(current_eviction_recode = case_when(str_detect(current_eviction_or_foreclosures, "doesn't know|refused|data not|n/a") ~ "unknown",
                                             str_detect(current_eviction_or_foreclosures, "none") ~ "no",
                                             is.na(current_eviction_or_foreclosures) ~ "unknown",
                                            TRUE ~ current_eviction_or_foreclosures ))
```

check for current eviction/foreclosure
```{r}
demo_dat %>% 
  count(current_eviction_or_foreclosures, current_eviction_recode)
```

Previous Evictions or Foreclosures - this one is truly so messy I will hold off until we see the new data with column issue fix 
```{r}
demo_dat %>% 
  count(previous_evictions_or_foreclosures)
```

Discharge from institution in the last 6 months - combining all unknown, leaving much of the rest for post column issue fix
```{r}
demo_dat <- demo_dat %>% 
  mutate(discharge_institution_recode = case_when(str_detect(discharge_from_institution_in_the_last_six_months, "data not|refused|doesn't know|n/a") ~ "unknown",
                                                  is.na(discharge_from_institution_in_the_last_six_months) ~ "unknown",
                                                  TRUE ~ discharge_from_institution_in_the_last_six_months))
```

Check for discharge from institution
```{r}
demo_dat %>% 
  count(discharge_from_institution_in_the_last_six_months, discharge_institution_recode)
```

APS report date - 

### Question (for Sara): How do I code that we will want to check that the date is something between start of the program and the present? 

Reporting Source - comining unknown responses into a "unknown" (including not applicable). Correcting spelling of "self", changing "client" to "self". Recoding "relative" and "family" as "family member".Changing "community professional" to "other community professional" to match their drop down. Changing "financial institution" and "financial" to "financial service provider". Combining mental health and mental health personnel.Changing "no relation" to "no relationship".

### Question:
  mandated reporter - ? (could be many, i.e. social worker, medical personnel, mh personnel)
  paid caregiver - another where I am unsure, maybe professional service provider?
  no APS referral, not applicable, no - (no might be column shift) but these, change to unknown? or is it interesting that some didn't get referred?
```{r}
demo_dat <- demo_dat %>% 
  mutate(reporting_source_recode = case_when(str_detect(reporting_source, "data not|refused|doesn't know|#|not applicable") ~ "unknown",
                                             str_detect(reporting_source, "sel|client") ~"self",
                                             str_detect(reporting_source, "relative|family") ~ "family member",
                                             str_detect(reporting_source, "community pro") ~"other community professional",
                                             str_detect(reporting_source, "financial") ~ "financial service provider",
                                             str_detect(reporting_source, "mental health") ~ "mental health personnel",
                                             str_detect(reporting_source, "no relatio") ~"no relationship",
                                             str_detect(reporting_source, "workshop|mandated report") ~ "flag",
                                             is.na(reporting_source) ~ "unknown",
                                             
                                             TRUE ~ reporting_source))
```

Check for reporting source
```{r}
demo_dat %>% 
  count(reporting_source, reporting_source_recode)
```

### Question: for previous APS involvement - seeing some responses that appear to fit in intervention, will this affect processing (given these are now subsets of data?) Hopefully this is just a column issue, but want to flag. 

### Question: anything we want to do with processing the income variables? (income fro benefits, work for pay, other income)





## To come back to if needed:

Clean APS Report Location indicator, remove "'",", ca" and " ca" from strings

```{r}
demo_dat <- demo_dat %>% 
  mutate(aps_report_location = str_replace(aps_report_location, "'", "")) %>% 
  mutate(aps_report_location = str_replace(aps_report_location, ", ca| ca|,ca", ""))
```

Examine APS Report Location

```{r}
demo_dat %>% 
  count(aps_report_location)
```

## Question for Gina

Should we drop everything from " -" (i.e. space and then dash)
```{r}
demo_dat %>%  
  count(aps_report_location) %>% 
  filter(str_detect(aps_report_location, "-")) %>% 
  filter(str_detect(aps_report_location, "-"))
```

replace all locations that contain Roseville in the name to just Roseville as a new location variable 
```{r}
demo_dat <- demo_dat %>% 
  mutate(aps_location_recode = case_when(str_detect(aps_report_location,"rosev") ~ "roseville",
                                      
                                         TRUE ~ aps_report_location))
```

check
```{r}
demo_dat %>% 
  count(aps_report_location, aps_location_recode) %>% 
  filter(str_starts(aps_report_location, "r"))
```

Counts of `gender`

```{r}
demo_dat %>% 
  count(race_1, gender_identity)
```

  - Ask Gina, is the data transposition between `race` variables and `gender` being corrected prior to getting to us?

What does the `gender` data look like when `race_1` equals`transgender` ,`female` or `male`

```{r}
demo_dat %>% 
  filter(race_1 %in% c("female", "male", "transgender")) %>% 
  count(gender_identity, race_1)
```

Was all the data with this issue just a matter of a column shift???

One solution (remove these--re-shift and then re add)

```{r}
shifted_data <- demo_dat %>% 
    filter(race_1 %in% c("female", "male", "transgender")) 

# Anti join, remove shifted data
demo_dat <- demo_dat %>% 
      filter(!race_1 %in% c("female", "male", "transgender")) 


# Fix the shifted data

shifted_data <- shifted_data %>% 
  mutate(
    location_of_participation = gender_identity,
    gender_identity = race_1,
    race_1 = race_2,
    race_2 = ethnicity
  )

```

Glimpse shifted data

```{r}
shifted_data %>% 
  count(gender_identity)

shifted_data %>% 
  count(race_1)

shifted_data %>% 
  count(ethnicity)
```

Looks good, now re-add the shifted data BACK to the main demo_dat.

```{r}
demo_dat <- demo_dat %>% 
  bind_rows(shifted_data)

# Remove the shifted data subset from RAM

rm(shifted_data)
```

Check data dimensions.

```{r}
demo_dat %>% 
  dim()
```

# Re-coding variables

```{r}


```
