---
title: "Cleaning Demographic Data"
author: "Sara Colom"
date: "`r Sys.Date()`"
output: 
  github_document:
    df_print: kable
---


# Read in data & libraries

```{r, Read in Data & Libraries, warning = F, message = F}
source("R/helpers.R")
```

```{r}
demo_dat %>% 
  glimpse()
```


# Clean up demographic data

Make all variables of `character` lower case, trim trailing and starting white spaces and 'extra' inner spaces.

```{r}
demo_dat <- demo_dat %>% 
  mutate(across(where(is.character), clean_characters))

```

Clean APS Report Location indicator, remove "'",", ca" and " ca" from strings

```{r}
demo_dat <- demo_dat %>% 
  mutate(aps_report_location = str_replace(aps_report_location, "'", "")) %>% 
  mutate(aps_report_location = str_replace(aps_report_location, ", ca| ca|,ca", ""))
```

Examine APS Report Location

```{r}
demo_dat %>% 
  count(aps_report_location)
```

## Question for Gina

Should we drop everything from " -" (i.e. space and then dash)
```{r}
demo_dat %>%  
  count(aps_report_location) %>% 
  filter(str_detect(aps_report_location, "-")) %>% 
  filter(str_detect(aps_report_location, "-"))
```

replace all locations that contain Roseville in the name to just Roseville as a new location variable 
```{r}
demo_dat <- demo_dat %>% 
  mutate(aps_location_recode = case_when(str_detect(aps_report_location,"rosev") ~ "roseville",
                                      
                                         TRUE ~ aps_report_location))
```

sanity check
```{r}
demo_dat %>% 
  count(aps_report_location, aps_location_recode) %>% 
  filter(str_starts(aps_report_location, "r"))
```


Counts of `gender`

```{r}
demo_dat %>% 
  count(race_1, gender_identity)
```

  - Ask Gina, is the data transposition between `race` variables and `gender` being corrected prior to getting to us?

What does the `gender` data look like when `race_1` equals`transgender` ,`female` or `male`

```{r}
demo_dat %>% 
  filter(race_1 %in% c("female", "male", "transgender")) %>% 
  count(gender_identity, race_1)
```

Was all the data with this issue just a matter of a column shift???

One solution (remove these--re-shift and then re add)

```{r}
shifted_data <- demo_dat %>% 
    filter(race_1 %in% c("female", "male", "transgender")) 

# Anti join, remove shifted data
demo_dat <- demo_dat %>% 
      filter(!race_1 %in% c("female", "male", "transgender")) 


# Fix the shifted data

shifted_data <- shifted_data %>% 
  mutate(
    location_of_participation = gender_identity,
    gender_identity = race_1,
    race_1 = race_2,
    race_2 = ethnicity
  )

```

Glimpse shifted data

```{r}
shifted_data %>% 
  count(gender_identity)

shifted_data %>% 
  count(race_1)

shifted_data %>% 
  count(ethnicity)
```

Looks good, now re-add the shifted data BACK to the main demo_dat.

```{r}
demo_dat <- demo_dat %>% 
  bind_rows(shifted_data)

# Remove the shifted data subset from RAM

rm(shifted_data)
```

Check data dimensions.

```{r}
demo_dat %>% 
  dim()
```

# Re-coding variables

```{r}


```
